---
title: "Annex to Econ homework"
output: html_notebook
---

Run required function from main book first.

Annex

Simulate a random walk processes $y_t = y_{t-1} + e_t, y_0 = 0, e_t \sim IID(0, \sigma^2)$ analog to (14.03)
```{r rw, echo=FALSE, message=FALSE, warning=FALSE, fig.align = 'center', out.width="50%"}
# Simulate : White noise, Random Walk
set.seed(102)
sigma_ux <- 1 ; 
T <- 300      # T = sample size (length of TS to simulate)
# Simulate white Noise: u_{xt} ~ iidN(0,1), for t = 1, ..., T   
u = rnorm(T+1, mean = 0, sd = sigma_ux);  
# Loop to simulate observations of Random Walk : X_t = X_{t-1} + u_{xt} , 
X.RW      <- NA*matrix(0,T+1,1) # empty vector ( T X 1 )
X.RW[1]   <- 0 # initial value of RW set arbitrarily = 0 (unconditional mean)

# 'for' loop to generate all observations of RW form 2 to T
for (t in 2:(T+1)){ X.RW[t] <- X.RW[t-1] + u[t] }
X.RW <- X.RW[-1,] # delete initial observation which was set to 0    

# plot simulated RW (X_t) and its innovations (u_xt)
# dev.off()
plot(
  X.RW, type='o', 
  pch=19, col="blue", cex = 0.5,
  main=TeX('White Noise and Random Walk'), 
  xlab='time', ylab='')
abline(0,0)
lines(u, type='o', pch=19, col="red", cex = 0.5)
legend(
  "bottomleft", pch=c(19,19), col=c("blue", "red"),
  legend=c(TeX('RW: $X_t = X_{t-1} + u_{x,t}$   '), TeX('WN: $u_{x,t} \\sim iidN(0,1)$   ')))

```

Simulate two AR(1) processes according to $x_t = \phi x_{t-1} + u_t$, with $\phi=0.5$ and with $\phi = 1$. The ACFs of two processes indicate that in latter significant autocorrelation persists at least up to 20 lags, while the former autocorrelation becomes insignificant after 3 lags. Hence, in the case of AR(1), autocorrelation is a function of the parameter $\phi$.

```{r ar, echo=FALSE, message=FALSE, warning=FALSE}
# Simulate AR(1) and plot acf
set.seed(102)
sigma_ux <- 1 ; 
T<- 200
u    <- rnorm(T+1, mean = 0, sd = sigma_ux) #WN
phi  <- 0.5                                 # AR 1 coefficient 

X.AR1.stat    <- NA*matrix(0,T,1) # empty vector ( T X 1 )
X.AR1.nonstat <- NA*matrix(0,T,1) # empty vector ( T X 1 )

X.AR1.nonstat[1] <- 0 # both processes start at t= 0 from the value 0
X.AR1.stat[1]    <- 0  
for (t in 2:T){
  X.AR1.nonstat[t] <- X.AR1.nonstat[t-1] + u[t];
  X.AR1.stat[t]    <- phi*X.AR1.stat[t-1] + u[t] 
}

y.range <-range(as.matrix(c(X.AR1.stat,X.AR1.nonstat)),na.rm = TRUE, finite = FALSE) 

par(mfrow=c(2,2))

plot(
  X.AR1.nonstat, 
  type='o', pch=19, col="blue", cex = 0.5,
  main=TeX('AR(1): $X_t = X_{t-1} + u_{x,t},  u_{x,t} \\sim iidN(0,1)$'), 
  xlab='time', ylab='')
abline(0,0)

plot(
  X.AR1.stat, 
  type='o', pch=19, col="red", cex = 0.5,
  main=TeX('AR(1): $X_t = 0.5 X_{t-1} + u_{x,t},  u_{x,t} \\sim iidN(0,1)$'), 
  xlab='time', ylab='')

# plot Autocorrelation function (acf)
lag.max.acf = 20;
acf(
  X.AR1.nonstat, 
  main=TeX('ACF, AR(1) NON-stationary'), 
  lag.max = lag.max.acf, xlab = "lag ")

acf(
  X.AR1.stat, 
  main=TeX('ACF, AR(1) stationary'), 
  lag.max = lag.max.acf, xlab = "lag ")
```

Define a function to produces N AR(1) simulations. 
```{r ar_function, echo=FALSE, message=FALSE, warning=FALSE}



phi_vec     = c(0.95, 0.95)
c_vec       = c(0,0)
sigma_u_vec = c(1,1) 
T           = 60

output_temp <- independent_AR_simulation(phi_vec, c_vec, sigma_u_vec, T)
AR_sim <- output_temp$AR_sim_mat
u_sim  <- output_temp$u_sim_mat
```

Plot the simulated error terms of the two AR(1)s, that is $u_{x,t}$ and $u_{y,t}$. The respective ACFs indicate that the error terms are not autocorrelated.

```{r acf_u_v, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,2))

plot(
  u_sim[,1], 
  type='o', pch=19, col="blue", cex = 0.5,
  main=TeX('AR innovation: $u_{x,t}$'), 
  xlab='time', ylab='')
abline(0,0)

plot(
  u_sim[,2], 
  type='o', pch=19, col="red", cex = 0.5,
  main=TeX('AR innovation: $u_{y,t}$'), 
  xlab='time', ylab='')
abline(0,0)

# ACFs
lag.max.acf = 30

acf(
  u_sim[,1], 
  main=TeX('ACF, AR: $x_t = \\phi_1 x_{t-1} + u_{x,t}$'), 
  lag.max = lag.max.acf, xlab = "lag ")

acf(
  u_sim[,2], 
  main=TeX('ACF, AR: $y_t = \\phi_2 y_{t-1} + u_{y,t}$'), 
  lag.max = lag.max.acf, xlab = "lag ")
```

Scatterplot of $u_{x,t}$ and $u_{y,t}$. The error terms, while individually and randomly generated, appear to to correlated, which is confirmed by the regression of $u_{y,t}$ on $u_{x,t}$. The resulting $t_\beta$ is significant.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width="50%"}
par(mfrow=c(1,1))
plot(x = u_sim[,1], y = u_sim[,2], 
     col = "blue", lwd =1, 
     xlab=TeX("$u_{x,t}$"), ylab=TeX("$u_{y,t}$"),
     main=TeX('$u_{y,t}$ vs. $u_{x,t}$'))
reg1  <- lm(u_sim[,2] ~ u_sim[,1])

abline(reg1, col="red", lwd=2)

stargazer(list(reg1),type="text")
summary(reg1)

```

Compare ACF of the two AR processes. Autocorrelation is present and takes a fading wave pattern.

```{r echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,2))

plot(
  AR_sim[,1], 
  type='o', pch=19, col="blue", cex = 0.5,
  main=TeX('AR: $x_t = \\phi_1 x_{t-1} + u_{x,t}$'), 
  xlab='time', ylab='')
abline(0,0)

plot(
  AR_sim[,2], 
  type='o', pch=19, col="red", cex = 0.5,
  main=TeX('AR: $y_t = \\phi_2 y_{t-1} + u_{y,t}$'), 
  xlab='time', ylab='')
abline(0,0)

lag.max.acf = 30
acf(
  AR_sim[,1], 
  main=TeX('ACF, AR: $x_t = \\phi_1 x_{t-1} + u_{x,t}$'), 
  lag.max = lag.max.acf, xlab = "lag ")

acf(
  AR_sim[,2], 
  main=TeX('ACF, AR: $y_t = \\phi_2 y_{t-1} + u_{y,t}$'), 
  lag.max = lag.max.acf, xlab = "lag ")

```

Run a regression of two AR(1) processes. 
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width="50%", fig.align='center'}
#################################################
# scatterplot of y_t vs. x_t
#################################################
par(mfrow=c(1,1))
plot(x = AR_sim[,1], y = AR_sim[,2], 
     col = "blue", lwd =1, 
     xlab=TeX("$x_{t}$"), ylab=TeX("$y_{t}$"),
     main=TeX('$y_{t}$ vs. $x_{t}$'),   )

reg2  <- lm(AR_sim[,2] ~ AR_sim[,1])
stargazer(list(reg2),type="text")
summary(reg2)
abline(reg2, col="red", lwd=2) # display regression line on scatterplot (y~x)
```

Compare the regression results. The errors as well as the AR(1) are correlated given significant $\beta$s
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width="50%", fig.align='center'}
stargazer(list(reg1,reg2), type = "text")
# Magics: you can save tabel directly in latex !!!!!!!!!!!!!!!!!!!
# stargazer(list(reg1,reg2),
#           align=TRUE, type = "text", no.space = TRUE,
#           title = "Table X", out = "fit.tex")
```

