---
title: "EDHEC PhD Finance 2022 - Econometrics Homework"
author: "Giovanni Maffei, Timo Predoehl"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

```{r set_up, message=TRUE, warning=TRUE, include=FALSE, paged.print=FALSE}
pkgs.installed <- installed.packages()
pkgs.required <- c("car", "AER", "stargazer", "tictoc", "quantmod", "xts", "readr", "latex2exp", "gridExtra", "summarytools", "qwraps2", "nortest", "moments", "xtable", "sm", "astsa", "tseries", "ggplot2", "ggplotgui", "shiny", "tidyverse")
pkgs.missing <- pkgs.required[which(!pkgs.required %in% pkgs.installed)]
lapply(pkgs.missing, install.packages, character.only = TRUE)
lapply(pkgs.required, require, character.only = TRUE)

rm(list=ls()) # clear all variables from enviroment/workspace
```


\section*{Question 1: Spurious Regressions (2 points)} 
\begin{itemize}
	\item \textbf{compulsory} ( a ) [1 Point]
	\begin{itemize}
		\item{i} Replicate the analysis leading to Figure 14.1 in Davidson MacKinnon (2005, book) by running 100,000 MC, or simply 10,000 MC if your computer is slow (instead of 1 million), and using $T = 6, 12, 60, 120, 240, 360, 480$. (i) Compute also for each sample size T the distribution of the $R^2s$ of the MC simulations with either 7 separate histograms, or one unique figure where you report on the $y-axis$ the 5\%, 10\% 25\%, 50\%, 75\%, 90\% and 95\% quantiles of the distributions of the simulated $R^2$, and on the $x-axis$ you have T = 6, 12, 60, 120, 240, 360, 480.
		\item{ii} Similarly (either with histograms, or with one plot of the quantiles) report the distributions of the estimates t-statistics for the test of the null H0 : $\beta_2$ = 0 and
		\item{iii} their empirical rejection frequencies (that is the empirical size of the tests),which is exactly the figure 14.1 in Davidson MacKinnon (2005, book).
	\end{itemize}
	\item ( b ) [1 Point] Based on the results obtained by answering to point (a) summarize the problems of spurious regressions in econometrics.
\end{itemize} 

\subsection*{(a)(i) Replication Davidson McKinnon, 2005, fig 14.1}

Simulate a random walk processes $y_t = y_{t-1} + e_t, y_0 = 0, e_t \sim IID(0, \sigma^2)$ analog to (14.03)
```{r rw, echo=FALSE, message=FALSE, warning=FALSE}
# Simulate : White noise, Random Walk
set.seed(102)
sigma_ux <- 1 ; 
T <- 300      # T = sample size (length of TS to simulate)
# Simulate white Noise: u_{xt} ~ iidN(0,1), for t = 1, ..., T   
u = rnorm(T+1, mean = 0, sd = sigma_ux);  
# Loop to simulate observations of Random Walk : X_t = X_{t-1} + u_{xt} , 
X.RW      <- NA*matrix(0,T+1,1) # empty vector ( T X 1 )
X.RW[1]   <- 0 # initial value of RW set arbitrarily = 0 (unconditional mean)

# 'for' loop to generate all observations of RW form 2 to T
for (t in 2:(T+1)){ X.RW[t] <- X.RW[t-1] + u[t] }
X.RW <- X.RW[-1,] # delete initial observation which was set to 0    

# plot simulated RW (X_t) and its innovations (u_xt)
# dev.off()
plot(
  X.RW, type='o', 
  pch=19, col="blue", cex = 0.5,
  main=TeX('White Noise and Random Walk'), 
  xlab='time', ylab='')
abline(0,0)
lines(u, type='o', pch=19, col="red", cex = 0.5)
legend(
  "bottomleft", pch=c(19,19), col=c("blue", "red"),
  legend=c(TeX('RW: $X_t = X_{t-1} + u_{x,t}$   '), TeX('WN: $u_{x,t} \\sim iidN(0,1)$   ')))

```

Simulate two AR(1) processes according to $x_t = \phi x_{t-1} + u_t$, with $\phi=0.5$ and with $\phi = 1$. The ACFs of two processes indicate that in latter significant autocorrelation persists at least up to 20 lags, while the former autocorrelation becomes insignificant after 3 lags. Hence, in the case of AR(1), autocorrelation is a function of the parameter $\phi$.

```{r ar, echo=FALSE, message=FALSE, warning=FALSE}
# Simulate AR(1) and plot acf
set.seed(102)
sigma_ux <- 1 ; 
T<- 200
u    <- rnorm(T+1, mean = 0, sd = sigma_ux) #WN
phi  <- 0.5                                 # AR 1 coefficient 

X.AR1.stat    <- NA*matrix(0,T,1) # empty vector ( T X 1 )
X.AR1.nonstat <- NA*matrix(0,T,1) # empty vector ( T X 1 )

X.AR1.nonstat[1] <- 0 # both processes start at t= 0 from the value 0
X.AR1.stat[1]    <- 0  
for (t in 2:T){
  X.AR1.nonstat[t] <- X.AR1.nonstat[t-1] + u[t];
  X.AR1.stat[t]    <- phi*X.AR1.stat[t-1] + u[t] 
}

y.range <-range(as.matrix(c(X.AR1.stat,X.AR1.nonstat)),na.rm = TRUE, finite = FALSE) 

par(mfrow=c(2,2))

plot(
  X.AR1.nonstat, 
  type='o', pch=19, col="blue", cex = 0.5,
  main=TeX('AR(1): $X_t = X_{t-1} + u_{x,t},  u_{x,t} \\sim iidN(0,1)$'), 
  xlab='time', ylab='')
abline(0,0)

plot(
  X.AR1.stat, 
  type='o', pch=19, col="red", cex = 0.5,
  main=TeX('AR(1): $X_t = 0.5 X_{t-1} + u_{x,t},  u_{x,t} \\sim iidN(0,1)$'), 
  xlab='time', ylab='')

# plot Autocorrelation function (acf)
lag.max.acf = 20;
acf(
  X.AR1.nonstat, 
  main=TeX('ACF, AR(1) NON-stationary'), 
  lag.max = lag.max.acf, xlab = "lag ")

acf(
  X.AR1.stat, 
  main=TeX('ACF, AR(1) stationary'), 
  lag.max = lag.max.acf, xlab = "lag ")
```

Define a function to produces N AR(1) simulations. 
```{r ar_function, echo=FALSE, message=FALSE, warning=FALSE}

independent_AR_simulation<- function(phi_vec, c_vec, sigma_u_vec, T){
  # Simulate two independent RW and run spurious regression
  
  # function to simulate K independent AR process with :
  # T       = n.f simulated dates (simulated sample size)
  # phi_vec = [\phi_1, ..., phi_N]' = autoregressive coefficients for each AR process
  # c_vec   = [c_1, ..., c_N] = constant in AR process
  # number of AR(1) processes to simulate
  N <- length(phi_vec) 
  
  # create empty matrix T X N that will contain all simulated AR processes
  AR_sim_mat <- NA*matrix(0,T+1,N) # empty vector ( (T+1) X 1 )
   u_sim_mat <- NA*matrix(0,T+1,N) # empty vector ( (T+1) X 1 )
  
  # loop over each AR process ind1
  for (ind1 in seq(1,N)) {
    #ind1 = 1
    c       <-       c_vec[ind1]
    phi     <-     phi_vec[ind1]
    sigma_u <- sigma_u_vec[ind1]

    # simulate vector of innovations zero mean, and sd given as input
    set.seed(100+ind1) # fix "seed" in order to produce same random numbers every time next line of code is called!
    u    <- rnorm(T+1, mean = 0, sd = sigma_u) #WN
    u_sim_mat[,ind1] <- u

    # generate starting observation for the process: 
    # unconditional mean (stationary processes) or 0 (for non-stat processes only!, as phi=1 and c/(1-phi)=NA)
    if (abs(phi)<1){
      AR_sim_mat[1,ind1] <- c/(1-phi) ; #start from unconditional value of the process
    } else{
      AR_sim_mat[1,ind1] <- 0 ; #start from unconditional value of the process
    }

    # simulate all other observations t=2 to T+1
    for (t in seq(2,T+1)) AR_sim_mat[t,ind1] <- phi*AR_sim_mat[t-1,ind1] + u[t]
  }

  # Output of function
  # 1 = simulated AR process
  # 2 = simulated innovations
  return(list(AR_sim_mat = AR_sim_mat, u_sim_mat = u_sim_mat))
}

phi_vec     = c(0.95, 0.95) 
c_vec       = c(0,0)
sigma_u_vec = c(1,1) 
T           = 60

output_temp <- independent_AR_simulation(phi_vec, c_vec, sigma_u_vec, T)
AR_sim <- output_temp$AR_sim_mat
u_sim  <- output_temp$u_sim_mat
```

Plot the simulated error terms of the two AR(1)s, that is $u_{x,t}$ and $u_{y,t}$. The respective ACFs indicate that the error terms are not autocorrelated.

```{r acf_u_v, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,2))

plot(
  u_sim[,1], 
  type='o', pch=19, col="blue", cex = 0.5,
  main=TeX('AR innovation: $u_{x,t}$'), 
  xlab='time', ylab='')
abline(0,0)

plot(
  u_sim[,2], 
  type='o', pch=19, col="red", cex = 0.5,
  main=TeX('AR innovation: $u_{y,t}$'), 
  xlab='time', ylab='')
abline(0,0)

# ACFs
lag.max.acf = 30

acf(
  u_sim[,1], 
  main=TeX('ACF, AR: $x_t = \\phi_1 x_{t-1} + u_{x,t}$'), 
  lag.max = lag.max.acf, xlab = "lag ")

acf(
  u_sim[,2], 
  main=TeX('ACF, AR: $y_t = \\phi_2 y_{t-1} + u_{y,t}$'), 
  lag.max = lag.max.acf, xlab = "lag ")
```

Scatterplot of $u_{x,t}$ and $u_{y,t}$. The error terms, while individually and randomly generated, appear to to correlated, which is confirmed by the regression of $u_{y,t}$ on $u_{x,t}$. The resulting $t_\beta$ is significant.

```{r}
par(mfrow=c(1,1))
plot(x = u_sim[,1], y = u_sim[,2], 
     col = "blue", lwd =1, 
     xlab=TeX("$u_{x,t}$"), ylab=TeX("$u_{y,t}$"),
     main=TeX('$u_{y,t}$ vs. $u_{x,t}$'))
reg1  <- lm(u_sim[,2] ~ u_sim[,1])
stargazer(list(reg1),type="text")
summary(reg1)
abline(reg1, col="red", lwd=2)
```

Compare ACF of the two AR processes. Autocorrelation is present and takes a fading wave pattern with significance at the initial 5-10 lags and then again around 15 lags.

```{r echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,2))

plot(
  AR_sim[,1], 
  type='o', pch=19, col="blue", cex = 0.5,
  main=TeX('AR: $x_t = \\phi_1 x_{t-1} + u_{x,t}$'), 
  xlab='time', ylab='')
abline(0,0)

plot(
  AR_sim[,2], 
  type='o', pch=19, col="red", cex = 0.5,
  main=TeX('AR: $y_t = \\phi_2 y_{t-1} + u_{y,t}$'), 
  xlab='time', ylab='')
abline(0,0)

lag.max.acf = 30
acf(
  AR_sim[,1], 
  main=TeX('ACF, AR: $x_t = \\phi_1 x_{t-1} + u_{x,t}$'), 
  lag.max = lag.max.acf, xlab = "lag ")

acf(
  AR_sim[,2], 
  main=TeX('ACF, AR: $y_t = \\phi_2 y_{t-1} + u_{y,t}$'), 
  lag.max = lag.max.acf, xlab = "lag ")

```

Run a regression of two AR(1) processes. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
#################################################
# scatterplot of y_t vs. x_t
#################################################
par(mfrow=c(1,1))
plot(x = AR_sim[,1], y = AR_sim[,2], 
     col = "blue", lwd =1, 
     xlab=TeX("$x_{t}$"), ylab=TeX("$y_{t}$"),
     main=TeX('$y_{t}$ vs. $x_{t}$'),   )

reg2  <- lm(AR_sim[,2] ~ AR_sim[,1])
stargazer(list(reg2),type="text")
summary(reg2)
abline(reg2, col="red", lwd=2) # display regression line on scatterplot (y~x)
```

Compare the regression results
```{r echo=FALSE, message=FALSE, warning=FALSE}
stargazer(list(reg1,reg2), type = "text")
# Magics: you can save tabel directly in latex !!!!!!!!!!!!!!!!!!!
# stargazer(list(reg1,reg2),
#           align=TRUE, type = "text", no.space = TRUE, 
#           title = "Table X", out = "texoutput/fit.tex")
```

